---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rosv

<!-- badges: start -->
[![R-CMD-check](https://github.com/al-obrien/rosv/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/al-obrien/rosv/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/al-obrien/rosv/branch/master/graph/badge.svg)](https://app.codecov.io/gh/al-obrien/rosv?branch=master)
<!-- badges: end -->

Use R to query the [Open Source Vulnerability (OSV) database](https://osv.dev/). This can be useful to 
cross check the database with packages used in `requirements.txt` (python) or `renv.lock` files (R).
The generated content can also be used to create block lists for curated repositories with Posit Package Manager. More details about the OSV project and associated API can be found here: [https://google.github.io/osv.dev/](https://google.github.io/osv.dev/)

## Installation

You can install {rosv} only from GitHub:
```{r, eval = FALSE}
remotes::install_github('al-obrien/rosv')
```


## Basic Examples

### Detect vulnerable packages

One of the simplest queries is to provide a package and ecosystem and return a `TRUE`/`FALSE` response 
informing you if that package has ever been listed with a vulnerability.

```{r}
library(rosv)

is_pkg_vulnerable(c('dask', 'dash'), ecosystem = c('PyPI', 'PyPI'), cache = TRUE)
```

### List package vulnerabilities

The most basic usage of {rosv} is to pull all versions of PyPI or CRAN packages listed 
on the OSV database using high-level functions such as `osv_query()` and `create_osv_list()`.

```{r, eval = FALSE}
# Query one package in PyPI for vulnerabilities
pkg_vul <- osv_query('dask', ecosystem = 'PyPI')
create_osv_list(pkg_vul)
```

```{r example, eval = FALSE}
# Pull the entire set of PyPI vulnerability data
pypi_vul <- create_osv_list(ecosystem = 'PyPI')
pypi_vul
```


## Use API helpers directly
Lower-level functionality is available to return more details about the API request and response 
contained within the R6 object. These are more flexible than their higher-level alternatives. 
By default, the results of these functions will be cached using {memoise}. This can be overriden 
by specifying `cache = FALSE`.
```{r, eval = FALSE}
# Returns entire response object to parse as you please.
osv_query_1('dask', ecosystem = 'PyPI')

# Returns the vulnerability IDs for packages in list
osv_querybatch('dask', ecosystem = 'PyPI')

# Return vulnerabilities from different ecosystems as vectors
osv_querybatch(c('dask', 'readxl'), ecosystem = c('PyPI', 'CRAN'))

# Grab details by vulns ID
osv_vulns('PYSEC-2021-387')

# Clear cache, if needed
clear_osv_cache()
```


## Creating a cross-referenced whitelist

When using a product such as {miniCRAN} or Posit Package Manager, there may be 
corporate requirements to limit what packages users can install. Although having a 
whitelist is often recommended, it should either specify the exact versions that are approved 
or exclude packages with known vulnerabilities. Given the amount of packages and versions, this 
is often difficult. The following method will take a vector of packages (from PyPI) and cross-reference against 
the OSV database. If packages are identified they are either entirely dropped, or the specific versions with flagged 
vulnerabilities are excluded.

```{r, eval = FALSE}
# List of packages of interest
python_pkg <- c('dask', 'tensorflow', 'keras')

# Create the xref whitelist
pypi_vul <- create_osv_list(as.data.frame = TRUE)
xref_pkg_list <- create_ppm_xref_whitelist(python_pkg, pypi_vul)

# Output requirements.txt which can be used with PPM product
writeLines(xref_pkg_list, 'requirements.txt')
```


## Development notes

{rosv} is still a young project. There are plans to extend its use. Currently it uses R6 
classes for its lower-level interface to the OSV API. Pagination functionality will be added 
once it is offered by {httr2}, which at time of writing is available but experimental.
There are also plans to have more types of returned details, such as returning just the version 
input for functions like `osv_query()`.
