---
title: "Introduction to rosv"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{rosv}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Package purpose

The {rosv} package provides two core purposes:

- Access information from the [Open Source Vulnerability (OSV) database](https://osv.dev/).
- Operate on vulnerability information and create formatted lists for package administration.

Consequently, the functions in {rosv} relate to querying or downloading content from OSV, parsing JSON content, and generating tables and lists regarding key information on package vulnerabilities. The OSV database is not specific to R related repositories such as CRAN; users can access information on any ecosystem available in OSV. For R users who also dabble in Python, they can search for package vulnerabilities within the PyPI repository while remaining in an R interface.

## Basic Examples

The following examples will outline an assortment of package functionality but first we must load the package!

```{r setup}
library(rosv)
```

### Detect vulnerable packages

One of the simplest queries is to provide a package and ecosystem and return a `TRUE`/`FALSE` response 
informing you if that package has ever been listed with a vulnerability.

```{r, eval = FALSE}
is_pkg_vulnerable(c('dask', 'dash'), ecosystem = c('PyPI', 'PyPI'))
```

### List package vulnerabilities

The most basic usage of {rosv} is to pull all versions of an ecosystem's packages (e.g. PyPI or CRAN) listed 
in the OSV database. This can be achieved using high-level functions such as `osv_query()` and `create_osv_list()`. 

```{r, eval = FALSE}
# Query one package in PyPI for vulnerabilities
pkg_vul <- osv_query('dask', ecosystem = 'PyPI')
create_osv_list(pkg_vul)
```

```{r example, eval = FALSE}
# Pull the entire set of PyPI vulnerability data
pypi_vul <- create_osv_list(ecosystem = 'PyPI')
pypi_vul
```


## Use API helpers directly

Lower-level functions return more detail about the API request and response 
contained within the R6 object. These are more flexible than their higher-level alternatives. 
By default, the results of these functions will be cached. This can be overriden 
by specifying `cache = FALSE`.

The higher-level API query function `osv_query()` aligns returned content type and parsing method, making it the preferred choice for the typical use-case.

```{r, eval = FALSE}
# Returns entire response object to parse as you please.
osv_query_1('dask', ecosystem = 'PyPI')

# Returns the vulnerability IDs for packages in list
osv_querybatch('dask', ecosystem = 'PyPI')

# Return vulnerabilities from different ecosystems as vectors
osv_querybatch(c('dask', 'readxl'), ecosystem = c('PyPI', 'CRAN'))

# Grab details by vulns ID
osv_vulns('PYSEC-2021-387')
```

## Result caching

By default, results from queries using API helpers (e.g. `osv_query()` or `osv_querybatch()`) will cache results using `memoise::memoise()`. The caching can be turned off directly using function parameters or globally reset using `clear_osv_cache()`. Caching is the default behavior to help enforce polite access of the OSV API.

```{r, eval = FALSE}
# Query without caching
osv_query('dask', ecosystem = 'PyPI', cache = FALSE)

# Clear cache, as needed
clear_osv_cache()
```


## Creating a cross-referenced whitelist

When using a product such as {miniCRAN} or Posit Package Manager there may be 
corporate requirements to limit what packages users can install. Although having a 
whitelist is often recommended, it should either specify the exact versions that are approved 
or exclude packages with known vulnerabilities. Given the sheer amount of packages and their versions, this 
is often difficult. The following method will take a vector of packages (from PyPI) and cross-reference against 
the OSV database. If packages are identified they are either entirely dropped, or the specific versions with flagged 
vulnerabilities are excluded.

```{r, eval = FALSE}
# List of packages of interest
python_pkg <- c('dask', 'tensorflow', 'keras')

# Create the xref whitelist
pypi_vul <- create_osv_list(as.data.frame = TRUE)
xref_pkg_list <- create_ppm_xref_whitelist(python_pkg, pypi_vul)

# Output requirements.txt which can be used with PPM product
writeLines(xref_pkg_list, file.path(tempdir(), 'requirements.txt'))
```
